02-preliminary-analysis

### 2024-05-24 data/02-preliminary-analysis.qmd
added 'access_15.rds', the parcel level cumulative opp count per parcel and category

```{r re-weighting points so all destinations are considered spatially equal in access counts, include=FALSE}
Dependent_Centric <- 0.25*2225/1265
Errand_Centric <- 0.25*2225/158
Grocery_Centric <- 0.25*2225/381
Health_Centric <- 0.25*2225/421

#adds up to 2225 opportunities  
0.25*2225/1265*(1265) +  0.25*2225/158*(158) + 0.25*2225/381*(381) + 0.25*2225/421*(421)

#adding the weighted care destination counts -- reweighted such that the 4 categories each make up 25% of the 2225 points.
care_dest <- care_dest |>
          mutate(Dest_count_weighted = case_when(Care_Category == "Dependent-centric" ~ Dependent_Centric,
                                                 Care_Category == "Errand-centric" ~ Errand_Centric,
                                                 Care_Category == "Grocery-centric" ~ Grocery_Centric,
                                                 Care_Category == "Health-centric" ~ Health_Centric))

care_dest$Dest_count_weighted |> sum()
```

```{r travel-time-15min}
#15 mins or less
ttm_care_walk_15 <- ttm_care_walk_parcel |> 
  filter(travel_time_p50 <= 15)|> 
  left_join(
    care_dest |> st_drop_geometry() |> select(c("care_ID", "Care_Category", "Care_Category_Specific", "Care_Category_Speciality", "Dest_count_weighted")),  by = c("to_id" = "care_ID")) |>
  mutate(f = 1)

@@

```{r}
access_15 <- ttm_care_walk_15 |> 
  group_by(from_id, Care_Category, Care_Category_Specific) |> 
  summarize(GeoUID = first(GeoUID),
            copp_isc = n(),
            copp_isc_w = sum(Dest_count_weighted))

test_all <- ttm_care_walk_15 |>
  sp_avail_detailed(o_id = from_id,
                    d_id = to_id, 
                    pop = pop_dwell, #pop assumed at each res. parcel -- it is the pop in the DA divided by the number of dwellings in the DA. 
                    opp = Dest_count_weighted, #instead of just being vector of 1
                    r = f, f = f, alpha = 1) |> 
  group_by(from_id, Care_Category, Care_Category_Specific) |> summarize(V_isc_w = sum(V_ij, na.rm = TRUE))

@@

access_15 <- bind_rows(access_15_zeros, access_15) |>
  group_by(from_id, Care_Category, Care_Category_Specific) |> 
  summarize(GeoUID = first(GeoUID),
            copp_isc = sum(copp_isc),
            copp_isc_w = sum(copp_isc_w),
            V_isc_w = sum(V_isc_w))
rm(access_15_zeros)
#check
access_15$V_isc_w |> sum(na.rm=T)
```

Aggregating time!

@@

#join this to access_15_bycategory_DA

access_15_bycategory_DA <- access_15_bycategory_DA |> left_join(step2, by="GeoUID")
```

```{r}
#the number of opportunities from each categories
all_cats <- access_15_bycategory_DA |> 
  mutate(copp_all = sum(copp_ic_med, na.rm=T) + sum(copp_ic_sd, na.rm=T),
         copp_all_w = sum(copp_ic_w_med, na.rm=T) + sum(copp_ic_w_sd, na.rm=T)) |>
  group_by(GeoUID) |>
         summarise(summed_copp_all = sum(copp_all),
                   summed_copp_all_w = sum(copp_all_w),
                   summed_Vic_w = sum(V_ic_w, na.rm=T),
                   Population = mean(Population),
                   Care_Access_Status1 = case_when(summed_copp_all < 4 ~ "[0-04 des.) Not caring",
                                 summed_copp_all >= 4 & summed_copp_all < 25 ~ "[04-25 des.) Caring", 
                                 summed_copp_all >= 25 ~ "[25+ des.) Very caring"),
                   Care_Access_Status1_w = case_when(summed_copp_all_w < 4 ~ "[0-04 des.) Not caring",
                                 summed_copp_all_w >= 4 & summed_copp_all_w < 25 ~ "[04-25 des.) Caring", 
                                 summed_copp_all_w >= 25 ~ "[25+ des.) Very caring"),

>>>>

```{r re-weighting points so all destinations are considered spatially equal in access counts, include=FALSE}
# Dependent_Centric <- 0.25*2225/1265
# Errand_Centric <- 0.25*2225/158
# Grocery_Centric <- 0.25*2225/381
# Health_Centric <- 0.25*2225/421
# 
# #adds up to 2225 opportunities  
# 0.25*2225/1265*(1265) +  0.25*2225/158*(158) + 0.25*2225/381*(381) + 0.25*2225/421*(421)
# 
# #adding the weighted care destination counts -- reweighted such that the 4 categories each make up 25% of the 2225 points.
# care_dest <- care_dest |>
#           mutate(Dest_count_weighted = case_when(Care_Category == "Dependent-centric" ~ Dependent_Centric,
#                                                  Care_Category == "Errand-centric" ~ Errand_Centric,
#                                                  Care_Category == "Grocery-centric" ~ Grocery_Centric,
#                                                  Care_Category == "Health-centric" ~ Health_Centric))
# 
# care_dest$Dest_count_weighted |> sum()
```

```{r travel-time-15min}
#15 mins or less
ttm_care_walk_15 <- ttm_care_walk_parcel |> 
  filter(travel_time_p50 <= 15)|> 
  left_join(
    care_dest |> st_drop_geometry() |> select(c("care_ID", "Care_Category", "Care_Category_Specific", "Care_Category_Speciality")),  by = c("to_id" = "care_ID")) |>
  mutate(f = 1)

@@

```{r}
access_15 <- ttm_care_walk_15 |> 
  group_by(from_id, Care_Category, Care_Category_Specific) |> 
  summarize(GeoUID = first(GeoUID),
            copp_isc = n())

test_all <- ttm_care_walk_15 |>
  sp_avail_detailed(o_id = from_id,
                    d_id = to_id, 
                    pop = pop_dwell, #pop assumed at each res. parcel -- it is the pop in the DA divided by the number of dwellings in the DA. 
                    opp = f, 
                    r = f, f = f, alpha = 1) |> 
  group_by(from_id, Care_Category, Care_Category_Specific) |> summarize(V_isc_w = sum(V_ij, na.rm = TRUE))

@@

access_15 <- bind_rows(access_15_zeros, access_15) |>
  group_by(from_id, Care_Category, Care_Category_Specific) |> 
  summarize(GeoUID = first(GeoUID),
            copp_isc = sum(copp_isc),
            V_isc_w = sum(V_isc_w))
rm(access_15_zeros)
#check
access_15$V_isc_w |> sum(na.rm=T)
```

Aggregating time!
```{r}
#ID = care destination ID (in care_dest), GeoUID is id of the DA, from_id is parcel ID
saveRDS(access_15, "access_15.rds")
```

@@

#join this to access_15_bycategory_DA
access_15_bycategory_DA <- access_15_bycategory_DA |> left_join(step2, by="GeoUID")
```
COMEBACKHERE
```{r}
#the number of opportunities from each categories
all_cats <- access_15_bycategory_DA |> 
  mutate(copp_all = sum(copp_ic_med, na.rm=T) + sum(copp_ic_sd, na.rm=T),
         copp_all_w = sum(copp_ic_w_med, na.rm=T) + sum(copp_ic_w_sd, na.rm=T)) |>
  group_by(GeoUID) |>
         summarise(summed_copp_all = sum(copp_all),
                   summed_copp_all_w = sum(copp_all_w),
                   summed_Vic_w = sum(V_ic_w, na.rm=T),
                   Population = mean(Population),
                   Care_Access_Status1 = case_when(summed_copp_all < 4 ~ "[0-04 des.) Not caring",
                                 summed_copp_all >= 4 & summed_copp_all < 25 ~ "[04-25 des.) Caring", 
                                 summed_copp_all >= 25 ~ "[25+ des.) Very caring"),
                   Care_Access_Status1_w = case_when(summed_copp_all_w < 4 ~ "[0-04 des.) Not caring",
                                 summed_copp_all_w >= 4 & summed_copp_all_w < 25 ~ "[04-25 des.) Caring", 
                                 summed_copp_all_w >= 25 ~ "[25+ des.) Very caring"),
